Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-24 Paper/Paper.tex
more small tweaks

\begin{figure}[thb]
\begin{center}
\includegraphics[width=.6\columnwidth]{code}
\end{center}
\caption{\small{A typical codeword from a codebook of size $200$ (code
    7 in Figure \ref{fig:codes1}), corresponding to a
    tonic-subdominant chord progression.  The patch is composed of $2$
    bars and the pattern length was set to $8$ beats.  }}
\label{fig:code}
\end{figure}

@@

The 25 most frequently used codewords in the test set are shown in
Figure \ref{fig:codes1}.  
The frequency of use of these codewords are shown in Figure \ref{fig:freqs}.
The codewords primarily consist of sustained
notes and simple chords.  Since they are designed to be key-invariant,
specific notes do not appear.  Instead the first 7 codewords
correspond to a single note sustained across two bars (codeword 0),
perfect fifth (codewords 1 and 2) and fourth intervals (codewords 3
and 6), and a
major and minor triads (codewords 5 and 4, respectively).  Many of the remaining
codewords correspond to common transitions from one chord to another,
\eg a tonic-subdominant transition in codes 7 and 9
(Cmaj
$\rightarrow$ Fmaj or C5 $\rightarrow$ F5 as a guitar power chord)
% assuming that chroma 0 is C) 
and the reverse tonic-dominant transition in code 21 (\ie Fmaj
$\rightarrow$ Cmaj).
%FIXME: Does this transition have a better name like ``relative x''?
% C5 -> F5 in 'guitar chords', basically a power chord

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes_lle}
\end{center}
\caption{\small{
Locally linear embedding (LLE) visualization of the codebook.
LLE tries to keep near neighbors (here according to euclidean distance)
in the same $2$D neighborhood. Shown patterns were randomly selected
in each neighborhood.
}}
\label{fig:lle}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{cc}
\toprule
Offset & \% of times chosen \\
\midrule
0 & $\mathbf{62.6}$\\
1 & $16.5$\\
2 & $09.4$\\
3 & $11.5$\\
\bottomrule
\end{tabular}
\end{center}
\caption{\small{
Song segmentation experiment:
best offset according to our codebook. The real one is $0$.
Random guess gives an accuracy of $25$\%.
}}
\label{tab:offset}
\end{table}

@@

All songs in the artist dataset are encoded as an histogram of the codes
used in that song. The frequency values are normalized by the number
of patterns in the song. We test each song in a leave-one-out setting.
Leaving aside the test song, we represent each of the $20$ artists by the 
average of
the song histograms of that artist. The test song is compared to
the histogram of each artist using euclidean distance.

@@

Looking at the most typical patterns per artist can be insightful.
To find these patterns, we compare its use by one artist and divide
by its use by all artists. In Figure \ref{fig:typicalpat} we see 
Metallica's pattern, then Tori Amos and
Suzanne Vega's pattern. These three artists were easily identified.
Artists like Metallica are characterized by ``wideband'' patterns -
featured muddled due to noisy acoustics - lots of transients,
distortion, etc, which can be seen in Figure \ref{fig:metallica2}.
%FIXME: ron, what di you mean: (not by cleaned up by EN's feature analysis)

@@

As for enhancements, we specifically did not compress our features using
Gaussian mixtures or other generative model. That being said, these methods
could be use to develop better distance measures between patterns.
The use of the euclidean distance is arbitrary. Summarizing patches
with Gaussians, and then comparing the distance between those Gaussians,
could reduce the influence of the noise in the distance measure.

Moving on to larger scales, we could imagine merging codes together and
splitting some on two if the clusters become too small or too large.We could
also cluster codebooks themselves, in a similar fashion as hierarchical
Gaussian mixtures \cite{Vasconcelos2001}.
%  (similar idea used for indexing \cite{Muja2009})

>>>>

\begin{figure}[thb]
\begin{center}
\includegraphics[width=.4\columnwidth]{code}
\end{center}
\caption{\small{A typical codeword from a codebook of size $200$ (code
    7 in Figure \ref{fig:codes1}), corresponding to a
    tonic-subdominant chord progression.  The patch is composed of $2$
    bars and the pattern length was set to $8$ beats.  }}
\label{fig:code}
\end{figure}

@@

The 25 most frequently used codewords in the test set are shown in
Figure \ref{fig:codes1}.  
The frequency of use of these codewords are shown in Figure \ref{fig:freqs}.
The codewords primarily consist of sustained
notes and simple chords.  Since they are designed to be key-invariant,
specific notes do not appear.  Instead the first 7 codewords
correspond to a single note sustained across two bars (codeword 0),
perfect fifth (codewords 1 and 2) and fourth intervals (codewords 3
and 6), and 
major and minor triads (codewords 5 and 4, respectively).  Many of the remaining
codewords correspond to common transitions from one chord to another,
\eg a tonic-subdominant transition in codes 7 and 9
(Cmaj
$\rightarrow$ Fmaj or C5 $\rightarrow$ F5 as a guitar power chord)
% assuming that chroma 0 is C) 
and the reverse tonic-dominant transition in code 21 (\ie Fmaj
$\rightarrow$ Cmaj).
%FIXME: Does this transition have a better name like ``relative x''?
% C5 -> F5 in 'guitar chords', basically a power chord

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes_lle}
\end{center}
\caption{\small{
Locally linear embedding (LLE) visualization of the codebook.
LLE tries to keep near neighbors (here according to Euclidean distance)
in the same $2$D neighborhood. Shown patterns were randomly selected
in each neighborhood.
}}
\label{fig:lle}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{cc}
\toprule
Offset & \% of times chosen \\
\midrule
0 & $\mathbf{62.6}$\\
1 & $16.5$\\
2 & $09.4$\\
3 & $11.5$\\
\bottomrule
\end{tabular}
\end{center}
\caption{\small{
Results from the song segmentation experiment.
The correct offset is $0$.
Baseline performance by choosing the offset randomly gives an accuracy of $25$\%.
}}
\label{tab:offset}
\end{table}

@@

All songs in the artist dataset are encoded as an histogram of the codes
used in that song. The frequency values are normalized by the number
of patterns in the song. We test each song in a leave-one-out setting.
Leaving aside the test song, we represent each of the $20$ artists by the 
average of
the song histograms of that artist. The test song is compared to
the histogram of each artist using Euclidean distance.

@@

Looking at the most typical patterns per artist can be insightful.
To find these patterns, we compare its use by one artist and divide
by its use by all artists. In Figure \ref{fig:typicalpat} we see 
Metallica's pattern, then Tori Amos and
Suzanne Vega's pattern. These three artists were easily identified.
Artists such as Metallica are characterized by ``wideband'' patterns,
containing features muddled by the noisy acoustics containing many
transients,
distortion, etc, which can be seen in Figure \ref{fig:metallica2}.
%FIXME: ron, what di you mean: (not by cleaned up by EN's feature analysis)

@@

As for enhancements, we specifically did not compress our features using
Gaussian mixtures or other generative model. That being said, these methods
could be use to develop better distance measures between patterns.
The use of the Euclidean distance is arbitrary. Summarizing patches
with Gaussians, and then comparing the distance between those Gaussians,
could reduce the influence of the noise in the distance measure.

Moving on to larger scales, we could imagine merging codes together and
splitting some on two if the clusters become too small or too large.  We could
also cluster codebooks themselves, in a similar fashion as hierarchical
Gaussian mixtures \cite{Vasconcelos2001}.
%  (similar idea used for indexing \cite{Muja2009})
